Fair, Stable Exchange Networks through Quenched Merchant Location and Idiosyncratic Trading Costs

### 2019-08-03 writing/main paper/main.tex
added first version of the short note. Still needs some proof reading, which I will do tomorrow.

Note that usually, it is not optimal for a firm to offer products so cheaply that it can sell to the entire market. 
It tends to be more profitable to sell a lower quantity locally at a higher price.\green{I don't know if I would go this far here.}

@@

\green{This is the third time we define ordinal distance. We should cut at least
one}
For any buyer $b$ and firm $f$, we define the ordinal distance $D = D(b,f)$
such that the firm $f$ that is physically closest to $b$ has distance
$D(b,f)=1$, the next closest has distance $D=2$, and the $n$th closest has
distance $D=n$.  Thus, if buyer $b$ buys from firm $f$, the total price,
including tax, is given by 
\begin{equation}
	p(b,f) = p_f + p_f ~ r~ \left( D(b,f)^\gamma - 1 \right)
\end{equation}
where $\gamma$ is the coefficient governing how quickly the tax increases, and
$r$ is a simple scaling coefficient (set to $100\%$ in the remainder of this
report\green{in teh bar plots, $r=1$}).

@@

\subsection{A Dynamic Economy}

In the previous subsection, we have demonstrated how a distance-tax-free economy disproportionally rewards essentially random, minor differences in firm competitiveness. 
But even more concerningly, the problem does not stay static: 
This disproportional revenue can be used to further invest in aggressive marketing, thereby increasing the difference between the two firms even more. 
This may ultimately run a firm out of business, despite initially being as competitive as the other firms.

We highlight this mechanism more quantitatively by running the following simple game:
The same economy as in subsection \ref{sec:static_economy} is considered (2 firms, 12 buyers, 6 closest buyers each). 
However, instead of just offering one static set of quantities and prices, here we consider multiple business cycles (e.g. quarters, or years). 
Firm $1$ starts with an initial production cost that is only $1\%$ below that of firm $2$'s. 
But then, each of the two firms use their profit to effectively subsidise their costs, allowing them to offer lower prices during the next round. 
The outcome of this experiment is depicted in figure \ref{fig:dynamic_game},

@@

\red{
mention
- multiple products
- fixed location
- implementation on digital market place
- vision
}

>>>>

Note that usually, it is not optimal for a firm to offer products so cheaply that it can sell to the entire market. 
It tends to be more profitable to sell a lower quantity locally at a higher price.

@@

For any buyer $b$ and firm $f$, we define the ordinal distance $D = D(b,f)$
such that the firm $f$ that is physically closest to $b$ has distance
$D(b,f)=1$, the next closest has distance $D=2$, and the $n$th closest has
distance $D=n$.  Thus, if buyer $b$ buys from firm $f$, the total price,
including tax, is given by 
\begin{equation}
	p(b,f) = p_f + p_f ~ r~ \left( D(b,f)^\gamma - 1 \right)
\end{equation}
where $\gamma$ is the coefficient governing how quickly the tax increases, and
$r$ is a simple scaling coefficient (set to $100\%$ in the remainder of this
report).

@@

\subsection{A Dynamic Economy} 
\label{sec:dynamic}

In the previous subsection, we have demonstrated how a no-transportation-cost economy disproportionally rewards essentially random, minor differences in firm competitiveness. 
But even more concerningly, the problem does not stay static: 
This disproportional revenue can be used to further under-cut prices, invest in aggressive marketing etc. 
This increases the difference between the two firms even more. 
This may ultimately run a firm out of business, despite initially being as competitive as the other firms.

We highlight this mechanism more quantitatively by running the following simple game:
The same economy as in subsection \ref{sec:static_economy} is considered (2 firms, 12 buyers, 6 closest buyers each). 
However, instead of just offering one static set of quantities and prices, here we consider multiple business cycles (e.g. quarters, or years). 
Firm $1$ starts with an initial production cost that is only $1\%$ below that of firm $2$'s. 
But then, each of the two firms use their profit to effectively subsidise their costs, proportional to the revenue they have received. 
This allows them to offer lower prices during the next round. 
The outcome of this experiment is depicted in Figure \ref{fig:dynamic_game} both with tax $(\gamma > 0)$ and without tax $(\gamma=0)$. 
\begin{figure}[!htb]
	\centering
	\includegraphics[width=\textwidth]{dynamic_game}
	\caption{	The left (right) column depicts a situation of repeated business cycles (`games'), without (with) a distance based idiosyncratic tax, $\gamma=0$ ($\gamma > 0$). 
			Initially, firm $1$ is able to produce only $1\%$ cheaper than firm $2$. 
			After each round, the firms use their revenue to subsidies their own prices (or alternatively: reinvest into higher production efficiency, etc.). 
			We witness in the left column that this essentially drives firm $2$ out of business, as it is is forced to decrease prices as low as firm $1$, in order to stay competitive. 
			Consequently,  firm $2$'s  profit decreases, until it would ultimately run out of business.
			If only a minimal distance-based tax is raised $(\gamma=0.1)$, such a scenario can be prevented: 
			Both firms are able to decrease their prices and remain competitive. 
			The dashed line in the lower left plot depicts the analytically calculable `Cournot price', i.e. the Nash-equilibrium price in a zero-tax-model.
			As we can see: the lines almost perfectly match, thus showing that our numerical results are consistent. 
			}
	\label{fig:dynamic_game}
\end{figure}
We witness that despite an initially marginal difference where firm $1$ is producing the products for only $1\%$ less than firm $2$, if we set $\gamma=0$. 
Firm $2$ ultimately runs out of business, as it is forced to compete with firm $1$. 
If $\gamma > 0$, both firms remain competitive. 
A detailed description of the outcome is found in the figure caption.

Although not shown here, we can deduce from row 5 in Figure \ref{fig:static_game} that even for $\gamma = 1/2$, firm $2$ is ran out of business, if the difference in cost is initially large, as anticipated. 
The only way firm $2$ can survive unconditionally of its competitiveness, is if $\gamma$ is set to effectively infinite. 
With basic calculus, we can calculate analytically an upper bound $\overline{\gamma}$ at which all firms are selling in effective monopolies (derivation will be shown in an Appendix of a next report).

@@

In this report, we have shown that problematic disproportional revenue is generated as our economy converges to a zero-transportation-cost limit. 
This is essentially rooted in the fact that initially minor (and essentially random) imbalances in product quality reward massively more the marginally cheaper/better product, 
as no rational agent has any incentive to deviate.

We have then introduced an idiosyncratic buyer-seller distance-based tax, that avoids such disproportional revenue in case the differences in production cost are minor, 
but at the same time continues to reward competitiveness in the sense of significantly cheaper production costs.

The current state of the paper serves as proof of concept, 
A lot of work has been devoted to determine the Nash equilibria correctly. 
Now that everything is set up, we can move forward quickly.

In particular, the following points should still be addressed: 
\red{@Sandy: please tell us what you think we should focus on}

For now, the tax revenue has not been used. 
We should implement a targeted redirection of the tax money. 
We have to be careful that we do not thereby distort the game-theoretical equilibria. 
So for instance, if we give back the tax revenue directly to the most effected firms (either the richest or the poorest), it essentially distorts their incentives. 
Funding of new start-ups, or an equal redistribution to customers comes to mind. 
\red{@Sandy: any practical suggestions?}

A value for $\gamma$ striking an optimal balance between competition and firm protection has to be set. 
A value too low leads to reward of essentially random, minor production advantages. 
A value too large dampens competition by protecting firms too much. 
We will have to find a balance, ideally one that is defined algorithmically and adjusts itself automatically. 
We can implement one $\gamma$ per product, and continue to consider a single-product economy. 
Ideally, we can couple it to the tax revenue, once that has been defined.

The requirement that firms and buyers are placed uniformly at random appears unnecessarily restrictive. 
If we want to impose such a mechanism e.g, on a digital market place such as Amazon, one should give buyers and sellers at least the illusion of having a choice. 
We can use our simulations to show that the model still holds, if buyers and sellers can chose their location on the circle freely, as long as change happens on time-scales much slower than multiple transactions.

Ultimately, we want to run more realistic simulations of real market places, such as Amazon.
